---
# title: ""
# author: "Mikhail Dozmorov"
# date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r setup, echo=FALSE, include=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Set up the environment
library(knitr) 
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', dpi=100,  echo=F, autodep=T, width=200, warning=FALSE, comment=NA)
options(replace.assign=TRUE, width=500)
tidy.opts=list(width.cutoff=200)
suppressMessages(library(pander))
panderOptions('table.split.table', Inf)
set.seed(1)
```

```{r message=FALSE}
library(xlsx)
# Load term mapping
term.mapping <- read.xlsx2("data/icd9_mapping.xlsx", sheetName="manual")
#term.mapping <- term.mapping[ term.mapping$Category == "immunological", ]
invisible(source("../../../../GenomeRunner/R.GenomeRunner/utils1.R")) # See the required packages there
source("../../../../GenomeRunner/R.GenomeRunner/episimilarity.R")
# Correlation to use
cortype <- "spearman"
```

Gene Epigenomic similarity
===

Load the data
```{r}
mtx <- load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_OR.txt", "data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt"))
# Subset by the names in term.mapping object
mtx <- mtx[, colnames(mtx) %in% term.mapping$Name]
```

Test which clusters are reproducible

```{r eval=FALSE}
library(pvclust)
library(parallel)
# cl <- makeCluster(4, type="PSOCK")
# result10K <- parPvclust(cl=cl, as.matrix(mtx), method.dist=dist.method, method.hclust=hclust.method, nboot=10000, weight=TRUE)
# saveRDS(result10K, "data/pvclust.Rds")
# stopCluster(cl)
result <- readRDS("data/pvclust.Rds")
plot(result, hang = -1)
pvrect(result, alpha=0.95)
seplot(result, identify=TRUE)
print(result, which=c(26))
```

Plotting heatmap of the original correlation matrix

```{r}
mtx.cor <- rcorr(mtx, type=cortype)[[1]]
# Reorder correlation matrix the same order as pvclust did
mtx.cor <- mtx.cor[ result$hclust$order, result$hclust$order]
# A vector of category names, same order as the correlation matrix
categoryNames <- term.mapping$Category[ match(rownames(mtx.cor), term.mapping$Name) ] # Reorder category names to match disease names orded
names(categoryNames) <- term.mapping$Name[ match(rownames(mtx.cor), term.mapping$Name)]
# Set side colors
ColSideColors <- as.numeric(factor(categoryNames[ colnames(mtx.cor) ]))
cbind((categoryNames), ColSideColors) # Sanity check
ColSideColors[ ColSideColors == 1 ] <- "red" # Hematological parameter
ColSideColors[ ColSideColors == 2 ] <- "blue" # Immunological
ColSideColors[ ColSideColors == 3 ] <- "green" # Metabolic
ColSideColors[ ColSideColors == 4 ] <- "yellow" # Neurological
# Plotting
  par(oma=c(5,0,0,5), mar=c(10, 4.1, 4.1, 5), cex.main=0.65) # Adjust margins
  my.breaks <- seq(min(mtx.cor[mtx.cor != min(mtx.cor)]), max(mtx.cor[mtx.cor != max(mtx.cor)]), length.out=(2*granularity + 1))
  h <- heatmap.2(as.matrix(mtx.cor), trace="none", density.info="none", col=color, cexRow=1, cexCol=1, breaks=my.breaks, main="Regulatory similarity clustering", RowSideColors=ColSideColors, ColSideColors=ColSideColors, Rowv=FALSE, Colv=FALSE)  
```

```{r eval=FALSE}
library("dynamicTreeCut")
tmp <- cutreeDynamic(dendro=as.hclust(mtx.heat$rowDendrogram), method="tree", minClusterSize=3, deepSplit=TRUE)
#tmp <- cutreeDynamic(dendro=as.hclust(mtx.heat$rowDendrogram), method="hybrid", distM=as.dist(mtx.heat$rowDendrogram), minClusterSize=10, deepSplit=0)
tmp
names(tmp) <- colnames(mtx.cor)[ mtx.heat$rowInd ]
#image(mtx.cor[ match(names(sort(tmp)), colnames(mtx.cor)), match(names(sort(tmp)), colnames(mtx.cor)) ])

tmp1 <- mtx.cor[ match(names(sort(tmp)), rownames(mtx.cor)), match(names(sort(tmp)), colnames(mtx.cor)) ]
categoryNames <- term.mapping$Category[ match(rownames(tmp1), term.mapping$Name) ]
names(categoryNames) <- term.mapping$Name[ match(rownames(tmp1), term.mapping$Name)]
# Set side colors
ColSideColors <- as.numeric(factor(categoryNames[ colnames(tmp1) ]))
cbind((categoryNames), ColSideColors) # Sanity check
ColSideColors[ ColSideColors == 1 ] <- "red" # Immunological
ColSideColors[ ColSideColors == 2 ] <- "blue" # Metabolic
ColSideColors[ ColSideColors == 3 ] <- "green" # Neurological
ColSideColors[ ColSideColors == 4 ] <- "yellow" # Trait
ColSideColors[ ColSideColors == 5 ] <- "black" # Trait

heatmap.2(tmp1, trace="none", density.info="none", col=color, Rowv= FALSE, Colv=FALSE, cexRow=1, cexCol=1, RowSideColors = ColSideColors)  

```

Example of testing combinations of clustering methods
```{r eval=FALSE}
dist.methods <- c("euclidean", "maximum")#,"manhattan", "minkowski")
hclust.methods <- c("ward.D", "ward.D2", "single", "complete", "average", "mcquitty", "median", "centroid")
pdf("results/clustering_combinations.pdf")
for (d in dist.methods) {
  for (h in hclust.methods) {
    par(oma=c(10,0,0,10), mar=c(10, 4.1, 4.1, 10), cex.main=0.65) # Adjust margins
    h<-heatmap.2(as.matrix(mtx.cor), trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=d)}, hclustfun=function(x){hclust(x, method=h)}, cexRow=1, cexCol=1,  main=paste(d, h),  RowSideColors=ColSideColors) # cellnote=formatC(as.matrix(mtx.cor), format="f", digits=2), notecol="black", notecex=1,
  }
}
dev.off()
```

Define clusters

```{r}
plot(result$hclust, hang = -1)
k <- 2 # Number of clusters. Use 5 for defining 4 clusters
rect.hclust(result$hclust, k=k)
clusters <- cutree(result$hclust, k=k)
# Remove clusters having less than or equal to 1 member
# clusters <- clusters[ clusters != as.data.frame(table(clusters), drop=F)$clusters[ as.data.frame(table(clusters), drop=F)$Freq <= 1 ] ]
mtx.clust <- list(eset.labels = names(clusters), eset.groups=as.vector(clusters))
```

```{r fig.height=3, eval=FALSE}
# Empirically define clusters
#par(mfrow=c(1, 2))
h <- 2 # Empirically set height to cut the tree
mtx.heat$colDendrogram %>% plot(horiz=F) #; abline(h=h)
rect.hclust(as.hclust(mtx.heat$rowDendrogram), h=h, border="red")
#as.hclust(mtx.heat$colDendrogram)$height %>% density %>% plot; abline(v=h)
mtx.clust <- mtx.heat$colDendrogram %>% mtx.clusters(height=h, minmembers=2)
#par(mfrow=c(1, 1))
```

Roadmap, Cell type-specific analysis
---

```{r}
mtx <- load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_PVAL.txt", "data.gr/Roadmap_DNase_hotspotbroadall/matrix_PVAL.txt", "data.gr/Roadmap_DNase_narrowPeak/matrix_PVAL.txt", "data.gr/Roadmap_gappedPeak/matrix_PVAL.txt", "data.gr/Roadmap_narrowPeak/matrix_PVAL.txt", "data.gr/Roadmap_chromStates18/matrix_PVAL.txt"))
mtx.cellspecific(mtx, "results/cellspecific_Roadmap.xlsx")
```

Encode, Cell type-specific analysis
---

```{r}
mtx <- load_gr_data(c("data.gr/ENCODE_Tfbs/matrix_PVAL.txt", "data.gr/ENCODE_Tfbs/matrix_PVAL.txt", "data.gr/ENCODE_Histone/matrix_PVAL.txt"))
mtx.cellspecific(mtx, "results/cellspecific_Encode.xlsx")
```

2 clusters, what are the DEGFs counts in each clustering?
===

broadPeak
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_broadPeak/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="broadPeak_2_clust"))
```

ENCODE TFBS
---

```{r}
mtx.degs <- load_gr_data("data.gr/ENCODE_Tfbs/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="encTfbs_2_clust"))
```

chromStates18
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_chromStates18/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="chromStates18_2_clust"))
```

DNase-hotspotbroadall
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="DNasehotspotbroadall_2_clust"))
```


Testing differences between autoimmune diseases
===
Define clusters

```{r}
plot(result$hclust, hang=-1)
k <- 5 # Number of clusters. Use 5 for defining 4 clusters
rect.hclust(result$hclust, k=k)
clusters <- cutree(result$hclust, k=k)
# Remove clusters having less than or equal to 1 member
clusters <- clusters[ clusters != as.data.frame(table(clusters))$clusters[ as.data.frame(table(clusters))$Freq <= 1 ] ]
mtx.clust <- list(eset.labels = names(clusters), eset.groups=as.vector(clusters))
```

broadPeak
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_broadPeak/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")

sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="broadPeak_4_clust"))
```

ENCODE TFBS
---

```{r}
mtx.degs <- load_gr_data("data.gr/ENCODE_Tfbs/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")

sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="encTfbs_4_clust"))
```

chromStates18
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_chromStates18/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")

sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="chromStates18_4_clust"))
```

DNase-hotspotbroadall
---

```{r}
mtx.degs <- load_gr_data("data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")

sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="DNasehotspotbroadall_4_clust"))
```



Exploring individual diseases
===

```{r}
# 1  Bone_mineral_density
# 2	Red_blood_cell_traits
# 3	Vitiligo
# 4	HDL_cholesterol
# 5	Platelet_counts
# 6	Primary_biliary_cirrhosis
# 7	Type_1_diabetes
# 8	Renal_function_related_traits_BUN
# 9	Alopecia_areata
# 10	Ankylosing_spondylitis
# 11	Multiple_sclerosis
# 12	Rheumatoid_arthritis
# 13	Kawasaki_disease
# 14	Liver_enzyme_levels_gamma_glutamyl_transferase
# 15	C_reactive_protein
# 16	LDL_cholesterol
# 17	Ulcerative_colitis
# 18	Restless_legs_syndrome
# 19	Urate_levels
# 20	Celiac_disease
# 21	Crohns_disease
# 22	Behcets_disease
# 23	Systemic_lupus_erythematosus
# 24	Atopic_dermatitis
# 25	Triglycerides
# 26	Psoriasis
# 27	Primary_sclerosing_cholangitis
# 28	Migraine
# 29	Allergy
# 30	Chronic_kidney_disease
# 31	Fasting_glucose_related_traits
# 32	Autoimmune_thyroiditis
# 33	Alzheimers_combined
# 34	Type_2_diabetes
# 35	Systemic_sclerosis
# 36	Asthma
# 37	Creatinine_levels
# 38	Progressive_supranuclear_palsy
# 39	Juvenile_idiopathic_arthritis
showHeatmap("data.gr/ENCODE_Tfbs/matrix_PVAL.txt", colnum=11, factor="none", cell="none", isLog10=FALSE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="barup", fileName=NULL)
showHeatmap("data.gr/ENCODE_Dnase/matrix_OR.txt", colnum=21, factor="none", cell="none", isLog10=TRUE, adjust="none", pval=1, numtofilt=1, toPlot="barup", fileName=NULL)

showHeatmap("data.gr/Roadmap_broadPeak/matrix_PVAL.txt", colnum=12, factor="none", cell=c("E033", "E034", "E037", "E038", "E039", "E040", "E041", "E042", "E043", "E044", "E045", "E047", "E048", "E062", "E115", "E116", "E123", "E124", "E029", "E030", "E031", "E032", "E035", "E036", "E046", "E050", "E051"), isLog10=FALSE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="barup", fileName=NULL)
showHeatmap("data.gr/Roadmap_chromStates18/matrix_PVAL.txt", colnum=12, factor="none", cell=c("E033", "E034", "E037", "E038", "E039", "E040", "E041", "E042", "E043", "E044", "E045", "E047", "E048", "E062", "E115", "E116", "E123", "E124", "E029", "E030", "E031", "E032", "E035", "E036", "E046", "E050", "E051"), isLog10=FALSE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="barup", fileName=NULL)

```

