---
# title: ""
# author: "Mikhail Dozmorov"
# date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r setup, echo=FALSE, include=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Set up the environment
library(knitr) 
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', dpi=100,  echo=F, autodep=T, width=200, warning=FALSE, comment=NA)
options(replace.assign=TRUE, width=500)
tidy.opts=list(width.cutoff=200)
suppressMessages(library(pander))
panderOptions('table.split.table', Inf)
set.seed(1)
# Local functions and settings
invisible(source("../../../../GenomeRunner/R.GenomeRunner/utils2.R")) # See the required packages there
source("../../../../GenomeRunner/R.GenomeRunner/episimilarity.R")
# Correlation to use
cortype <- "spearman"
```

Load the data
===

```{r message=FALSE, results='hide'}
library(xlsx)
# Load term mapping
term.mapping <- read.xlsx2("data/icd9_mapping.xlsx", sheetName="manual")
# term.mapping <- term.mapping[ term.mapping$BED != "Kawasaki_disease", ] # Exclude Kawasaki outlier
# Load the actual episimilarity data
mtx <- as.matrix(load_gr_data(c("data.gr/2Roadmap_broadPeak/matrix_OR.txt", "data.gr/2Roadmap_DNase_hotspotbroadall/matrix_OR.txt")))
mtx <- mtx[, match(term.mapping$BED, colnames(mtx))] # Match and subset the data
# A vector of category names, same order as the correlation matrix
categoryNames <- term.mapping$Category[ match(colnames(mtx), term.mapping$Name) ]
names(categoryNames) <- term.mapping$Name[ match(colnames(mtx), term.mapping$Name)]
# Set side colors
ColSideColors <- as.numeric(factor(categoryNames[ colnames(mtx) ]))
cbind((categoryNames), ColSideColors) # Sanity check
ColSideColors[ ColSideColors == 1 ] <- "red" # Immunological
ColSideColors[ ColSideColors == 2 ] <- "blue" # Metabolic
ColSideColors[ ColSideColors == 3 ] <- "green" # Neurological
ColSideColors[ ColSideColors == 4 ] <- "yellow" # Trait
```

We first check how the distributions of the log2-transformed odds ratios for each sample look like side-by-side.

```{r eval=TRUE, results='hold'}
ggplot(melt(mtx), aes(x=Var2, y=value, fill=Var2)) + 
  geom_boxplot() + 
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  theme(legend.position="none")
#   coord_cartesian(ylim = c(-5, 5)) +
```

It's a good idea to scale and center the data. Let's check how the scaled data looks like.

```{r results='hold'}
mtx <- scale(mtx)
ggplot(melt(mtx), aes(x=Var2, y=value, fill=Var2)) + 
  geom_boxplot() + 
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1)) +
  theme(legend.position="none")
#   coord_cartesian(ylim = c(-5, 5)) +
```

Testing number of clusters with COMMUNAL
===

```{r eval=FALSE}
# Testing cluster number with COMMUNAL
library(COMMUNAL) 
ks <- seq(2,6)
result <- COMMUNAL(data=mtx, ks=ks, dist.metric = "correlation")
result.list <- list(list(result), ncol(mtx))
goodAlgs <- c("hierarchical", "kmeans")
goodMeasures <- c('wb.ratio', 'avg.silwidth', 'dunn')
values <- plotRange3D(result.list, ks, goodAlgs, goodMeasures)
# Testing cluster number with ConsensysClusterPlus
library(ConsensusClusterPlus)
results = ConsensusClusterPlus(mtx, maxK=6, reps=50, pItem=0.8, pFeature=1, title="ConsensusClusterPlus", clusterAlg="hc", distance="spearman", seed=1262118388.71279, plot="png")
# Store hclust object
hclustergram <- results[[3]][["consensusTree"]]
```

Prepare the matrix of correlation coefficients, and cluster it.

```{r}
## Using spearman correlations
# rcorr returns a list, [[1]] - correl coeffs, [[3]] - p-values. Type - pearson/spearman
mtx.cor <- rcorr(as.matrix(mtx), type="spearman")[[1]]
# Optionally, try Spearman or Kendall correlation
# mtx.cor[[1]] <- cor(as.matrix(mtx), method="kendall")
# Or, MIC
# library(minerva) # Playing with MIC correlation coefficient
# mtx.cor <- mine(mtx, n.cores=4)$MIC
# Or, using distance metric
# "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski"
# dist.method <- "euclidean"
# mtx.cor <- dist(t(mtx), method = dist.method) # Get distance matrix

## Clustering
# "ward.D", "ward.D2", "single", "complete", "average" (= UPGMA), "mcquitty" (= WPGMA), "median" (= WPGMC) or "centroid" (= UPGMC)
hclust.method <- "ward.D2"
hclustergram <- hclust(as.dist(1-mtx.cor), method=hclust.method) # Cluster it
```

Gene Epigenomic similarity
===

```{r fig.height=6}
h <- mtx.plot(mtx.cor, SideColors=ColSideColors) # Clustering parameters
```

Plotting with aheatmap
---

```{r fig.height=6}
library(NMF)
annot <- data.frame(Category=term.mapping$Category, Size=as.numeric(term.mapping$BEDcount))
annotColor <- list(Category=c("red", "blue", "green", "yellow"), Size=c("red", "white", "blue"))
h <- aheatmap(mtx.cor, Rowv=as.dendrogram(hclustergram), Colv=as.dendrogram(hclustergram), color=matlab.like(50), distfun=dist.method, hclustfun=hclust.method, annCol=annot, annColors = annotColor)#, cexCol = 1, cexRow = 1)
```

Testing clustering stability

```{r eval=FALSE}
library(pvclust)
library(parallel)
cl <- makeCluster(4, type="PSOCK")
set.seed(1)
result <- parPvclust(cl=cl, mtx, method.dist=dist.method, method.hclust=hclust.method, nboot=1000)
stopCluster(cl)
# saveRDS(result, "data/pvClust_10000.Rds")
# result <- readRDS("data/pvClust_10000.Rds")
dev.off()
plot(result, hang = -1)
pvrect(result, alpha=0.95)
hclustergram <- result$hclust # Make the heatmap plot whatever hclust object
```

It appears scaled data produces less stable clustering, but more visually appealing. For now, skip pvclust. Visualizing what pvClust clustered.

```{r epigenomicVisualization1, fig.height=6, eval=FALSE}
h <- heatmap.2(mtx.cor, symkey = TRUE, Rowv=as.dendrogram(hclustergram), Colv=as.dendrogram(hclustergram), dendrogram="both", trace="none", density.info="none", col=color, RowSideColors=ColSideColors, ColSideColors=ColSideColors)  
```

Example of testing combinations of clustering methods
```{r eval=FALSE}
dist.methods <- c("euclidean", "maximum","manhattan", "minkowski")
hclust.methods <- c("ward.D", "ward.D2", "single", "complete", "average", "mcquitty", "median", "centroid")
pdf("results/clustering_combinations.pdf")
for (d in dist.methods) {
  for (h in hclust.methods) {
    par(oma=c(10,0,0,10), mar=c(10, 4.1, 4.1, 10), cex.main=0.65) # Adjust margins
    h<-heatmap.2(as.matrix(mtx.cor), trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=d)}, hclustfun=function(x){hclust(x, method=h)}, cexRow=1, cexCol=1,  main=paste(d, h),  RowSideColors=ColSideColors) # cellnote=formatC(as.matrix(mtx.cor), format="f", digits=2), notecol="black", notecex=1,
  }
}
dev.off()
```

```{r fig.height=3, eval=FALSE}
# Empirically define clusters
par(mfrow=c(1, 2), cex=0.5)
plot(as.dendrogram(h$Colv, hang=-1)) # Plot dendrogram
cl_num <- 3 # Empirically set desired numter of clusters
cols <- rainbow(cl_num) # Make colors for clusters
rect.hclust(as.hclust(h$Colv), k=cl_num, border=cols) # Define the clusters by rectangles
hcut <- 2# Empirically set height to cut the tree
abline(h=hcut)
as.hclust(h$Colv)$height %>% density %>% plot
abline(v=hcut)
mtx.clust <- h$Colv %>% mtx.clusters(height=hcut, minmembers=3)
par(mfrow=c(1, 1))
# Save the results of clustering
write.table(as.data.frame(mtx.clust), "results/clustering_all.txt", sep="\t", row.names=FALSE, quote=FALSE)
```

What are the DEGFs counts in each clustering?
===

chromStates18
---

```{r eval=FALSE}
mtx.degs <- load_gr_data("data.gr/2Roadmap_chromStates18/matrix_OR.txt")
mtx.degs <- mtx.degs[, match(term.mapping$BED, colnames(mtx))]
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="chromStates18_2_clust"))
```

DNase-hotspotbroadall
---

```{r eval=FALSE}
mtx.degs <- load_gr_data("data.gr/2Roadmap_DNase_hotspotbroadall/matrix_OR.txt")
mtx.degs <- mtx.degs[, match(term.mapping$BED, colnames(mtx))]
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="DNasehotspotbroadall_2_clust"))
```

broadPeak
---

```{r eval=FALSE}
mtx.degs <- load_gr_data("data.gr/2Roadmap_broadPeak/matrix_OR.txt")
mtx.degs <- mtx.degs[, match(term.mapping$BED, colnames(mtx))] 
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="broadPeak_2_clust"))
```

ENCODE TFBS
---

```{r eval=FALSE}
mtx.degs <- load_gr_data("data.gr/2ENCODE_Tfbs/matrix_OR.txt")
mtx.degs <- mtx.degs[, match(term.mapping$BED, colnames(mtx))]
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="encTfbs_2_clust"))
```


Testing random chance, Using GENES OR chromStates18+encHistone+encBroadHmm
---

```{r eval=FALSE}
mtx.degs <- load_gr_data(c("data.gr/2Roadmap_broadPeak/matrix_OR.txt", "data.gr/2Roadmap_DNase_hotspotbroadall/matrix_OR.txt", "data.gr/2ENCODE_Tfbs/matrix_OR.txt"))
mtx.degs <- mtx.degs[, match(term.mapping$BED, colnames(mtx))]
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust))
# Random simulations
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "row")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "col")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "mix")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "rnd")
sum(mtx.degfs(mtx.rnd[, mtx.clust$eset.labels], mtx.clust))
```

Cell type-specific analysis
---

```{r eval=FALSE}
mtx <- load_gr_data(c("data.gr/2Roadmap_broadPeak/matrix_PVAL.txt", "data.gr/2Roadmap_DNase_hotspotbroadall/matrix_PVAL.txt", "data.gr/2Roadmap_DNase_narrowPeak/matrix_PVAL.txt", "data.gr/2Roadmap_gappedPeak/matrix_PVAL.txt", "data.gr/2Roadmap_narrowPeak/matrix_PVAL.txt", "data.gr/2Roadmap_chromStates18/matrix_PVAL.txt"))
mtx.cellspecific(mtx, "results/cellspecific_Roadmap.xlsx")
```


