---
# title: ""
# author: "Mikhail Dozmorov"
# date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r setup, echo=FALSE, include=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
# Set up the environment
library(knitr) 
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', dpi=100,  echo=F, autodep=T, width=200, warning=FALSE, comment=NA)
options(replace.assign=TRUE, width=500)
tidy.opts=list(width.cutoff=200)
suppressMessages(library(pander))
panderOptions('table.split.table', Inf)
set.seed(1)
```

```{r message=FALSE}
library(xlsx)
# Load term mapping
term.mapping <- read.xlsx2("data/icd9_mapping.xlsx", sheetName="manual")
#term.mapping <- term.mapping[ term.mapping$Category == "immunological", ]
invisible(source("../../../../GenomeRunner/R.GenomeRunner/utils1.R")) # See the required packages there
source("../../../../GenomeRunner/R.GenomeRunner/episimilarity.R")
# Correlation to use
cortype <- "pearson"
```

Gene Epigenomic similarity
===

```{r}
# mtx.fig1a <- as.matrix(read.table("data/fig1a_matrix", sep="\t", header=F, stringsAsFactors=F))
# labels.fig1a <- readLines("data/fig1a_labels")
# colnames(mtx.fig1a) <- labels.fig1a; rownames(mtx.fig1a) <- labels.fig1a
# mtx.cor <- rcorr(mtx.fig1a, type="pearson")[[1]]

mtx.cor <- rcorr(load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_PVAL.txt")), type="pearson")[[1]]
# library(minerva) # Playing with MIC correlation coefficient
# mtx.cor <- mine(load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_PVAL.txt")), n.cores=4)$MIC
                
# Subset by the names in term.mapping object
mtx.cor <- mtx.cor[rownames(mtx.cor) %in% term.mapping$Name, colnames(mtx.cor) %in% term.mapping$Name]

# A vector of category names, same order as the correlation matrix
categoryNames <- term.mapping$Category[ match(rownames(mtx.cor), term.mapping$Name) ]
names(categoryNames) <- term.mapping$Name[ match(rownames(mtx.cor), term.mapping$Name)]
# Set side colors
ColSideColors <- as.numeric(factor(categoryNames[ colnames(mtx.cor) ]))
cbind((categoryNames), ColSideColors) # Sanity check
ColSideColors[ ColSideColors == 1 ] <- "red" # Immunological
ColSideColors[ ColSideColors == 2 ] <- "blue" # Metabolic
ColSideColors[ ColSideColors == 3 ] <- "green" # Neurological
ColSideColors[ ColSideColors == 4 ] <- "yellow" # Trait
mtx.heat <- mtx.plot(mtx.cor, SideColors=ColSideColors) # Clustering parameters
```

Testing clustering stability

```{r eval=FALSE}
mtx <- as.matrix(load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_OR.txt", "data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt")))
mtx.cor <- rcorr(mtx, type=cortype)[[1]]
dev.off()
library(pvclust)
library(parallel)
cl <- makeCluster(4, type="PSOCK")
result <- parPvclust(cl=cl, mtx, method.dist=dist.method, method.hclust=hclust.method, nboot=1000)
#result <- parPvclust(cl=cl, lung, method.dist="cor", method.hclust="average", nboot=10000)
stopCluster(cl)
plot(result)
pvrect(result, alpha=0.95)
```

Visualizing what pvClust clustered

```{r epigenomicVisualization1, fig.height=5}
# A vector of category names, same order as the correlation matrix
categoryNames <- term.mapping$Category[ match(colnames(mtx.cor), term.mapping$Name) ]
names(categoryNames) <- term.mapping$Name[ match(colnames(mtx.cor), term.mapping$Name)]
# Set side colors
ColSideColors <- as.numeric(factor(categoryNames[ colnames(mtx.cor) ]))
(cbind(categoryNames, ColSideColors)) # Sanity check
ColSideColors[ ColSideColors == 1 ] <- "red" # Immunological
ColSideColors[ ColSideColors == 2 ] <- "blue" # Metabolic
ColSideColors[ ColSideColors == 3 ] <- "green" # Neurological
ColSideColors[ ColSideColors == 4 ] <- "yellow" # Trait
h <- heatmap.2(mtx.cor, symkey = TRUE, Rowv=as.dendrogram(result$hclust), Colv=as.dendrogram(result$hclust), dendrogram="both", trace="none", density.info="none", col=color, RowSideColors=ColSideColors, ColSideColors=ColSideColors)  
```

Example of testing combinations of clustering methods
```{r eval=FALSE}
dist.methods <- c("euclidean", "maximum","manhattan", "minkowski")
hclust.methods <- c("ward.D", "ward.D2", "single", "complete", "average", "mcquitty", "median", "centroid")
pdf("results/clustering_combinations.pdf")
for (d in dist.methods) {
  for (h in hclust.methods) {
    par(oma=c(10,0,0,10), mar=c(10, 4.1, 4.1, 10), cex.main=0.65) # Adjust margins
    h<-heatmap.2(as.matrix(mtx.cor), trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=d)}, hclustfun=function(x){hclust(x, method=h)}, cexRow=1, cexCol=1,  main=paste(d, h),  RowSideColors=ColSideColors) # cellnote=formatC(as.matrix(mtx.cor), format="f", digits=2), notecol="black", notecex=1,
  }
}
dev.off()
```

```{r fig.height=3}
# Empirically define clusters
par(mfrow=c(1, 2))
hcut <- 30 # Empirically set height to cut the tree
h$colDendrogram %>% plot(horiz=T)
abline(v=hcut)
as.hclust(h$colDendrogram)$height %>% density %>% plot
abline(v=hcut)
mtx.clust <- h$colDendrogram %>% mtx.clusters(height=hcut, minmembers=2)
par(mfrow=c(1, 1))
```

What are the DEGFs counts in each clustering?
===

chromStates18
---

```{r}
# A matrix to use for differential epigenomic analysis
mtx.degs <- load_gr_data("data.gr/Roadmap_chromStates18/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="chromStates18"))
```

DNase-hotspotbroadall
---

```{r}
# A matrix to use for differential epigenomic analysis
mtx.degs <- load_gr_data("data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="DNasehotspotbroadall"))
```

broadPeak
---

```{r}
# A matrix to use for differential epigenomic analysis
mtx.degs <- load_gr_data("data.gr/Roadmap_broadPeak/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="broadPeak"))
```

ENCODE TFBS
---

```{r}
# A matrix to use for differential epigenomic analysis
mtx.degs <- load_gr_data("data.gr/ENCODE_Tfbs/matrix_OR.txt")
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter(var.cutoff=0.1)# %>% mtx.rand(randomize = "mix")
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust, label="encTfbs"))
```


Testing random chance, Using GENES OR chromStates18+encHistone+encBroadHmm
---

```{r}
# A matrix to use for differential epigenomic analysis
mtx.degs <- load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_OR.txt", "data.gr/Roadmap_DNase_hotspotbroadall/matrix_OR.txt", "data.gr/ENCODE_Tfbs/matrix_OR.txt"))
mtx.degs <- mtx.degs[, colnames(mtx.cor) %in% term.mapping$Name] %>% varFilter
sum(mtx.degfs(mtx.degs[, mtx.clust$eset.labels], mtx.clust))
# Random simulations
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "row")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "col")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "mix")
mtx.rnd <- mtx.degs %>% mtx.rand(randomize = "rnd")
sum(mtx.degfs(mtx.rnd[, mtx.clust$eset.labels], mtx.clust))
```

Cell type-specific analysis
---

```{r}
mtx <- load_gr_data(c("data.gr/Roadmap_broadPeak/matrix_PVAL.txt", "data.gr/Roadmap_DNase_hotspotbroadall/matrix_PVAL.txt", "data.gr/Roadmap_DNase_narrowPeak/matrix_PVAL.txt", "data.gr/Roadmap_gappedPeak/matrix_PVAL.txt", "data.gr/Roadmap_narrowPeak/matrix_PVAL.txt", "data.gr/Roadmap_chromStates18/matrix_PVAL.txt"))

mtx.cellspecific(mtx, "results/cellspecific_Roadmap.xlsx")

```


