---
title: "Comparison of manually classified diseases"
author: "Mikhail G. Dozmorov"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, echo=FALSE, message=FALSE}
library(MDmisc)
#source("/Users/mikhail/Documents/Work/GenomeRunner/R.GenomeRunner/utils2.R")
#source("/Users/mikhail/Documents/Work/GenomeRunner/R.GenomeRunner/episimilarity.R")
# Set up the environment
col3 <- colorRampPalette(c("blue", "yellow", "red"))
library(knitr) 
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F) #out.width=700, 
options(replace.assign=TRUE, stringsAsFactors = FALSE)#, width=120)
suppressMessages(library(pander))
panderOptions('table.split.table', Inf)
set.seed(1)
```

```{r localFunctions}
#' Reshape wide format data
#'
#' A function to reshape wide data format removing self-self and duplicated reverse associations
#'
#' @param mtx a matrix, in wide format. First column is the 'id.var'
#'
#' @return a matrix in long format, with self-self and duplicate associations removed
#' @export
#' @examples
#' reshape_wide_data(mtx)
##
reshape_wide_data <- function(mtx) {
  mtx <- mtx[ order(rownames(mtx)), order(colnames(mtx))]
  mtx[ lower.tri(mtx, diag=TRUE) ] <- NaN # Set lower portion to NaN, including the diagonal
  mtx <- reshape2::melt(mtx) # Melt, and remove Nan associations
  mtx <- mtx[ !is.nan(mtx$value), ]
  rownames(mtx) <- paste(mtx$Var1, mtx$Var2, sep="_")
  mtx <- mtx[, 3, drop=F]
  return(mtx)
}

library(xlsx)
# Load term mapping
term.mapping <- read.xlsx2("data/icd9_mapping.xlsx", sheetName="manual")
term.mapping <- term.mapping[term.mapping$DOID != "", ]

# Correlation to use
cortype <- "spearman" # 'sparman' or 'pearson'
mat.type <- "PVAL" # "PVAL" or "OR", specify whether matrix_PVAL or matrix_OR should be used 
```

**Current settings: Correlation: `r paste(cortype)`, Similarity measure: `r paste(mat.type)`**

Preparing Barabasi similarity

```{r diseaseNetwork1, echo=FALSE, cache=FALSE, eval=TRUE}
# All disease-disease relationships
mtx.disease <- readRDS("/Users/mikhail/Documents/Work/GenomeRunner/Paper-Similarity/data_gwasCatalog/AllNet5/AllNet5.Rds") # Full Barabasi matrix
# Optionally, use 3-digits IDC9 codes
# mtx.disease <- read.table("../../AllNet3.net", sep="\t", header=F)
# Subset by ICD9 codes associated with terms
mtx.disease <- mtx.disease[ mtx.disease$V1 %in% term.mapping$ICD9[ term.mapping$ICD9 != "" ] & mtx.disease$V2 %in% term.mapping$ICD9[ term.mapping$ICD9 != ""], ]

# Create empty matrix of term disease-disease relationships ($V5 - co-occurrence)
term.ICD9.occur <- matrix(NA, nrow=nrow(term.mapping), ncol=nrow(term.mapping))
rownames(term.ICD9.occur) <- term.mapping$Name
colnames(term.ICD9.occur) <- term.mapping$Name
# Create two others, to hold relative risk ($V6) and phi-correlation ($V9)
term.ICD9.risk <- term.ICD9.occur; term.ICD9.phi <- term.ICD9.occur

# Populate this matrix with Barabasi relationship values
for (i in 1:nrow(term.mapping)) {
  for (j in 1:nrow(term.mapping)) {
    if(term.mapping$ICD9[i] != "" & term.mapping$ICD9[j] != "") {
      # The disease associations matrix is not symmetrical - we pull up the index with either disease1-disease2 relationship, or disease2-disease1 relationship
      idx <- (mtx.disease$V1 == term.mapping$ICD9[i] & mtx.disease$V2 == term.mapping$ICD9[j]) | (mtx.disease$V2 == term.mapping$ICD9[i] & mtx.disease$V1 == term.mapping$ICD9[j])
      # If relationship pair is found, store it. Self-self relationships will be 0
      term.ICD9.occur[i, j] <- max(mtx.disease$V5[idx])
      term.ICD9.risk[i, j] <- max(mtx.disease$V6[idx])
      term.ICD9.phi[i, j] <- max(mtx.disease$V9[idx])
    }
  }
  # If self-self relationship is -Inf, replace it with maximum existing
  term.ICD9.occur[i, i] <- max(term.ICD9.occur[i, ], na.rm=T)
  term.ICD9.risk[i, i] <- max(term.ICD9.risk[i, ], na.rm=T)
  term.ICD9.phi[i, i] <- max(term.ICD9.phi[i, ], na.rm=T)
}
# Replace infinite by NAs
term.ICD9.occur[ is.infinite(term.ICD9.occur) ] <- NA
term.ICD9.phi[ is.infinite(term.ICD9.phi) ] <- NA
term.ICD9.risk[ is.infinite(term.ICD9.risk) ] <- NA
# Reformat the data
term.ICD9.occur <- reshape_wide_data(term.ICD9.occur)
term.ICD9.phi <- reshape_wide_data(term.ICD9.phi)
term.ICD9.risk <- reshape_wide_data(term.ICD9.risk)
# # Save the data
# saveRDS(term.ICD9.occur, "data/term.ICD9.3.occur.Rds")
# saveRDS(term.ICD9.phi, "data/term.ICD9.3.phi.Rds")
# saveRDS(term.ICD9.risk, "data/term.ICD9.3.risk.Rds")
# # Read in saved objects, ICD9 5-digits
# term.ICD9.occur <- readRDS("data/term.ICD9.occur.Rds")
# term.ICD9.phi <- readRDS("data/term.ICD9.phi.Rds")
# term.ICD9.risk <- readRDS("data/term.ICD9.risk.Rds")
# # Read in saved objects, ICD9 3-digits
# term.ICD9.occur <- readRDS("data/term.ICD9.3.occur.Rds")
# term.ICD9.phi <- readRDS("data/term.ICD9.3.phi.Rds")
# term.ICD9.risk <- readRDS("data/term.ICD9.3.risk.Rds")
```

Preparing AR similarity

```{r}
mtx.ar <- read.table("data/ar_11-11-2015.txt", sep="\t", header = T, as.is = T)
```


```{r eval=FALSE}
# Preparing Zhou similarity
mtx.sig <- read.table("/Users/mikhail/Documents/Work/GenomeRunner/gwas2bed/human-sig-disease-network.txt.gz", sep="\t", fill=T, header=T, as.is=T)
# Subset by ZhouIDs associated with our terms
mtx.sig <- mtx.sig[ mtx.sig$MeSH.Disease.Term %in% term.mapping$ZhouID[ term.mapping$ZhouID != "" ] & mtx.sig$MeSH.Disease.Term.1 %in% term.mapping$ZhouID[ term.mapping$ZhouID != ""], ]

# Create empty matrix of term disease-disease relationships ($V5 - co-occurrence)
term.Zhou.sig <- matrix(NA, nrow=nrow(term.mapping), ncol=nrow(term.mapping))
rownames(term.Zhou.sig) <- term.mapping$Name
colnames(term.Zhou.sig) <- term.mapping$Name

# Populate this matrix with Barabasi relationship values
for (i in 1:nrow(term.mapping)) {
  for (j in 1:nrow(term.mapping)) {
    if(term.mapping$ZhouID[i] != "" & term.mapping$ZhouID[j] != "") {
      # The disease associations matrix is not symmetrical - we pull up the index with either disease1-disease2 relationship, or disease2-disease1 relationship
      idx <- (mtx.sig$MeSH.Disease.Term == term.mapping$ZhouID[i] & mtx.sig$MeSH.Disease.Term.1 == term.mapping$ZhouID[j]) | (mtx.sig$MeSH.Disease.Term.1 == term.mapping$ZhouID[i] & mtx.sig$MeSH.Disease.Term == term.mapping$ZhouID[j])
      # If relationship pair is found, store it. Self-self relationships will be 0
      term.Zhou.sig[i, j] <- max(as.numeric(mtx.sig$symptom.similarity.score[idx]), na.rm=T)
    }
  }
  # If self-self relationship is -Inf, replace it with maximum existing
  term.Zhou.sig[i, i] <- max(as.numeric(term.Zhou.sig[i, ]), na.rm=T)
}
# Replace infinite by NAs
term.Zhou.sig[ is.infinite(term.Zhou.sig) ] <- NA
# Reformat the data
term.Zhou.sig <- reshape_wide_data(term.Zhou.sig)
```

```{r eval=FALSE}
# Preparing DiseaseConnect similarity 
mtx.sig <- read.table("/Users/mikhail/Documents/Work/GenomeRunner/gwas2bed/Disease-Disease_v1.txt", sep="\t", fill=T, header=T, as.is=T)
# Subset by DiseaseConnectIDs associated with our terms
mtx.sig <- mtx.sig[ mtx.sig$Disease.1 %in% term.mapping$DiseaseConnectID[ term.mapping$DiseaseConnectID != "" ] & mtx.sig$Disease.2 %in% term.mapping$DiseaseConnectID[ term.mapping$DiseaseConnectID != ""], ]

# Create empty matrix of term disease-disease relationships
term.DC.sig <- matrix(NA, nrow=nrow(term.mapping), ncol=nrow(term.mapping))
rownames(term.DC.sig) <- term.mapping$Name
colnames(term.DC.sig) <- term.mapping$Name

# Populate this matrix with Barabasi relationship values
for (i in 1:nrow(term.mapping)) {
  for (j in 1:nrow(term.mapping)) {
    if(term.mapping$DiseaseConnectID[i] != "" & term.mapping$DiseaseConnectID[j] != "") {
      # The disease associations matrix is not symmetrical - we pull up the index with either disease1-disease2 relationship, or disease2-disease1 relationship
      idx <- (mtx.sig$Disease.1 == term.mapping$DiseaseConnectID[i] & mtx.sig$Disease.2 == term.mapping$DiseaseConnectID[j]) | (mtx.sig$Disease.2 == term.mapping$DiseaseConnectID[i] & mtx.sig$Disease.1 == term.mapping$DiseaseConnectID[j])
      # If relationship pair is found, store it. Self-self relationships will be 0
      # We use minimum, because it's a p-value
      term.DC.sig[i, j] <- min(as.numeric(mtx.sig$P.value[idx]), na.rm=T)
    }
  }
  # If self-self relationship is -Inf, replace it with minimum existing
  term.DC.sig[i, i] <- min(as.numeric(term.DC.sig[i, ]), na.rm=T)
}
# Replace infinite by NAs
term.DC.sig[ is.infinite(term.DC.sig) ] <- NA
# -log10-transform the p-values, note that 0 p-values will be inf. Manually replace them with max values
tmp <- gr_transform(term.DC.sig)
tmp.max <- reshape2::melt(tmp)$value
tmp.max <- max(tmp.max[!(is.infinite(tmp.max)) & !(is.na(tmp.max))])
tmp[is.infinite(tmp)] <- tmp.max
# Reformat the data
term.DC.sig <- reshape_wide_data(tmp)
```

Preparing disease ontology similarity

```{r}
library(DOSE)
# # Validating some associations manually
# xx <- as.list(DOTERM)
# xx.term <- Term(names(xx))
# xx.term[ grepl("testicular", xx.term) ]
# xx.term[ grepl("testicular", xx.term, ignore.case=TRUE) & grepl("tumor", xx.term, ignore.case=TRUE)]
# xx.term["DOID:5419"]

dose.select <- function(measure="Wang"){
  tmp <- doSim(term.mapping$DOID[ term.mapping$DOID != "" ], term.mapping$DOID[ term.mapping$DOID != "" ], measure = measure) # "Wang", "Resnik", "Rel", "Jiang", and "Lin"
  rownames(tmp) <- term.mapping$Name[ term.mapping$DOID != "" ]
  colnames(tmp) <- term.mapping$Name[!(term.mapping$DOID == "")]
  return(reshape_wide_data(tmp))
}
dose.wang <- dose.select("Wang")
dose.resnik <- dose.select("Resnik")
dose.rel <- dose.select("Rel")
dose.jiang <- dose.select("Jiang")
dose.lin <- dose.select("Lin")
```

```{r eval=FALSE}
# Loading genomic jaccard similarity
mtx <- as.matrix(read.table("data/jaccard.autoimmune.wide.txt", sep="\t", header=T, row.names=1, stringsAsFactors=F))
mtx <- mtx[rownames(mtx) %in% term.mapping$Name, colnames(mtx) %in% term.mapping$Name] 
mtx.jaccard <- reshape_wide_data(mtx)
```

Loading Farh clustering

```{r}
mtx.fig1a <- as.matrix(read.table("data/fig1a_matrix", sep="\t", header=F, stringsAsFactors=F))
labels.fig1a <- readLines("data/fig1a_labels")
colnames(mtx.fig1a) <- labels.fig1a
rownames(mtx.fig1a) <- labels.fig1a
mtx.fig1a <- mtx.fig1a[term.mapping$Name, term.mapping$Name] # Subset by Names that are mapped
mtx.fig1a <- reshape_wide_data(mtx.fig1a)
```

```{r eval=FALSE}
# Loading GenometriCorr similarity
# mtx.gtc <- as.matrix(read.table("data/gtc.reldist.ecdfdeviation.wide.txt", sep="\t", header=T, row.names=1, stringsAsFactors=F))
# mtx.gtc <- as.matrix(read.table("data/gtc.scaled.absolute.min.distance.sum.1.wide.txt", sep="\t", header=T, row.names=1, stringsAsFactors=F))
mtx.gtc <- as.matrix(read.table("data/gtc.scaled.absolute.min.distance.sum.1.wide.txt", sep="\t", header=T, row.names=1, stringsAsFactors=F))
mtx.gtc <- mtx.gtc[rownames(mtx.gtc) %in% term.mapping$Name, colnames(mtx.gtc) %in% term.mapping$Name] 
mtx.gtc <- reshape_wide_data(mtx.gtc)
```

```{r eval=FALSE}
# Shrunken matrix
mtx.shrunken <- as.matrix(read.table("data/Similarity_Shrunken.txt", sep="\t", header=T, row.names=1, stringsAsFactors = FALSE))
mtx.shrunken <- mtx.shrunken[rownames(mtx.shrunken) %in% term.mapping$Name, colnames(mtx.shrunken) %in% term.mapping$Name] 
mtx.shrunken <- reshape_wide_data(mtx.shrunken)
```

Preparing literature similarity

```{r}
read.lit <- function(fname){
  mtx <- read.table(fname, sep="\t", stringsAsFactors = FALSE)
  mtx <- mtx[ mtx$V1 %in% term.mapping$name | mtx$V2 %in% term.mapping$Name, ]  # Subset by Names that are mapped
  tmp <- reshape2::dcast(mtx, V1 ~ V2, value.var="V3", mean)
  rownames(tmp) <- tmp$V1
  tmp$V1 <- NULL
  return(reshape_wide_data(as.matrix(tmp)))
}
# term.msh.path <- read.lit("data/terms.msh.path.txt")
# term.snomed.path <- read.lit("data/terms.snomed.path.txt")
# term.snomed.vector <- read.lit("data/terms.snomed.vector.txt")
term.iridescent.minMim <- read.lit("data/iridescent.minMim.txt")
term.iridescent.misn <- read.lit("data/iridescent.misn.txt")
term.iridescent.obsExp <- read.lit("data/iridescent.obsExp.txt")
term.iridescent.relOverlap <- read.lit("data/iridescent.relOverlap.txt")
term.iridescent.sharedRels <- read.lit("data/iridescent.sharedRels.txt")
```

Combining with episimilarity

```{r}
# Merge previously prepared matrixes
mtx.list.epi <- c(list(mtx.fig1a), list(term.iridescent.minMim), list(term.iridescent.misn), list(term.iridescent.obsExp), list(term.iridescent.relOverlap), list(term.iridescent.sharedRels), list(dose.jiang), list(dose.lin), list(dose.rel), list(dose.resnik), list(dose.wang)) 
#, list(mtx.jaccard)
#, list(term.DC.sig)
#, list(term.ICD9.occur), list(term.ICD9.risk), list(term.ICD9.phi) 
#, list(mtx.gtc), list(mtx.shrunken), list(mtx.clusterSim.MF.Resnik), list(mtx.clusterSim.MF.Wang))
#, list(term.msh.path), list(term.snomed.path), list(term.snomed.vector)

mtx.cor.epi.joined <- do.call(cbind, lapply(mtx.list.epi, function(x) x[match(rownames(mtx.list.epi[[1]]), rownames(x)), ]))
rownames(mtx.cor.epi.joined) <- rownames(mtx.list.epi[[1]])

# Set the column names after merging lists
colnames(mtx.cor.epi.joined) <- c("Shared.loci", "Iridescent.minMim", "Iridescent.misn", "Iridescent.obsExp", "Iridescent.relOverlap", "Iridescent.sharedRels", "DOSE.jiang", "DOSE.lin", "DOSE.rel", "DOSE.resnik", "DOSE.wang") 
#, "DiseaseConnect"
#, "ICD9.occur", "ICD9.risk", "ICD9.phi", "genometricorr", "shrunken")
#  "Jaccard", "MSH.path", "SnoMed.path", "SnoMed.vector",
```

Adding GenomeRunner data

```{r}
# Add (combinations of) GenomeRunner's results
if (mat.type == "PVAL") {
  dname <- list("data.gr.05/grweb_encBroadHmm//matrix_PVAL.txt",
                "data.gr.05/grweb_encChromCombined//matrix_PVAL.txt",
                "data.gr.05/grweb_encChromHmm//matrix_PVAL.txt",
                "data.gr.05/grweb_encSegway//matrix_PVAL.txt",
                "data.gr.05/grweb_encDNase//matrix_PVAL.txt",
                "data.gr.05/grweb_encHistone//matrix_PVAL.txt",
                "data.gr.05/grweb_encTFBS_cellspecific//matrix_PVAL.txt",
                "data.gr.05/grweb_encTFBS_clustered//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapChrom15//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapChrom18//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapChrom25//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_PVAL.txt",
                "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_PVAL.txt",
             c("data.gr.05/grweb_encBroadHmm//matrix_PVAL.txt", "data.gr.05/grweb_encChromCombined//matrix_PVAL.txt", "data.gr.05/grweb_encChromHmm//matrix_PVAL.txt", "data.gr.05/grweb_encDNase//matrix_PVAL.txt", "data.gr.05/grweb_encHistone//matrix_PVAL.txt", "data.gr.05/grweb_encSegway//matrix_PVAL.txt", "data.gr.05/grweb_encTFBS_cellspecific//matrix_PVAL.txt", "data.gr.05/grweb_encTFBS_clustered//matrix_PVAL.txt", "data.gr.05/grweb_roadmapChrom15//matrix_PVAL.txt", "data.gr.05/grweb_roadmapChrom18//matrix_PVAL.txt", "data.gr.05/grweb_roadmapChrom25//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_PVAL.txt"),
             c("data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_PVAL.txt"),
             c("data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_PVAL.txt", "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_PVAL.txt")
           )
} else if (mat.type == "OR") {
  dname <- list("data.gr.05/grweb_encBroadHmm//matrix_OR.txt",
                "data.gr.05/grweb_encChromCombined//matrix_OR.txt",
                "data.gr.05/grweb_encChromHmm//matrix_OR.txt",
                "data.gr.05/grweb_encSegway//matrix_OR.txt",
                "data.gr.05/grweb_encDNase//matrix_OR.txt",
                "data.gr.05/grweb_encHistone//matrix_OR.txt",
                "data.gr.05/grweb_encTFBS_cellspecific//matrix_OR.txt",
                "data.gr.05/grweb_encTFBS_clustered//matrix_OR.txt",
                "data.gr.05/grweb_roadmapChrom15//matrix_OR.txt",
                "data.gr.05/grweb_roadmapChrom18//matrix_OR.txt",
                "data.gr.05/grweb_roadmapChrom25//matrix_OR.txt",
                "data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_OR.txt",
                "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_OR.txt",
             c("data.gr.05/grweb_encBroadHmm//matrix_OR.txt", "data.gr.05/grweb_encChromCombined//matrix_OR.txt", "data.gr.05/grweb_encChromHmm//matrix_OR.txt", "data.gr.05/grweb_encDNase//matrix_OR.txt", "data.gr.05/grweb_encHistone//matrix_OR.txt", "data.gr.05/grweb_encSegway//matrix_OR.txt", "data.gr.05/grweb_encTFBS_cellspecific//matrix_OR.txt", "data.gr.05/grweb_encTFBS_clustered//matrix_OR.txt", "data.gr.05/grweb_roadmapChrom15//matrix_OR.txt", "data.gr.05/grweb_roadmapChrom18//matrix_OR.txt", "data.gr.05/grweb_roadmapChrom25//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_OR.txt"),
             c("data.gr.05/grweb_roadmapDNase_bPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_nPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_bPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_gPk-processed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_nPk-processed//matrix_OR.txt"),
             c("data.gr.05/grweb_roadmapDNase_gPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapDNase_nPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_gPk-imputed//matrix_OR.txt", "data.gr.05/grweb_roadmapHistone_nPk-imputed//matrix_OR.txt")
           )
}

# Read in data in a loop
for (x in dname) {
  mtx <- gr_load_data(x, p2z = T)
  mtx <- mtx[, colnames(mtx) %in% term.mapping$Name]  # Subset by Names that are mapped
  mtx.cor <- Hmisc::rcorr(mtx, type=cortype)[[1]]
  # mtx.cor <- mine(mtx, n.cores=4)$MIC
  mtx.cor.long <- reshape_wide_data(mtx.cor)
  mtx.cor.epi.joined <- cbind(mtx.cor.epi.joined, mtx.cor.long[match(rownames(mtx.cor.epi.joined), rownames(mtx.cor.long)), ])
}

# Set the column names after merging lists
colnames(mtx.cor.epi.joined)[ (length(mtx.list.epi) + 1):length(colnames(mtx.cor.epi.joined))] <- c("enc.BroadHmm", "enc.ChromCombined", "enc.ChromHmm", "enc.Segway", "enc.DNase", "enc.Histone", "enc.TFBS.cellspecific", "enc.TFBS.clustered", "rdm.Chrom15", "rdm.Chrom18", "rdm.Chrom25", "rdm.DNase.bPk.processed", "rdm.DNase.gPk.imputed", "rdm.DNase.nPk.imputed", "rdm.DNase.nPk.processed", "rdm.Histone.bPk.processed", "rdm.Histone.gPk.imputed", "rdm.Histone.gPk.processed", "rdm.Histone.nPk.imputed", "rdm.Histone.nPk.processed", "Combined.ALL", "Processed.ALL", "Imputed.ALL")
```

Adding GoShifter results

```{r}
# Function to load GoShifter data
gs_load_data <- function(postfix) {
  mtx.list <- list()
  for (i in term.mapping$BED) {
    mtx.list <- c(mtx.list, list(as.matrix(read.table(
      paste("data.gosifter/", i, "-gf_", postfix, ".txt", sep=""), 
      sep="\t", header = F, row.names = 1, stringsAsFactors = F, check.names = F), drop = F)))
  }
  
  mtx <- do.call("cbind", lapply(mtx.list, function(x) x[match(rownames(mtx.list[[1]]), rownames(x)), , drop = F]))
  class(mtx) <- "numeric"
  colnames(mtx) <- term.mapping$BED
  mtx <- gr_transform(mtx, p2z = F)
  mtx[is.infinite(mtx)] <- max(mtx[!is.infinite(mtx)])
  # Trim the matrix
  mtx <- mtx[apply(mtx, 1, function(x) sum(!is.na(x))) > 0, apply(mtx, 2, function(x) sum(!is.na(x))) > 
               0, drop = FALSE]  # Remove rows/columns with all NAs
  mtx <- mtx[!(apply(mtx, 1, function(x) sum(x == 0) == ncol(mtx))), , drop = F]  # If all values in a row are 0, remove these rows
  return(mtx)
}
# All GF categories
postfixes <- c("encchromStates", "encDNase", "encHistone", "encTFBS_cellspecific", "encTFBS_clustered", "rdmchromStates15", "rdmchromStates18", "rdmchromStates25", "rdmDNase_bPk-processed", "rdmDNase_gPk-imputed", "rdmDNase_nPk-imputed", "rdmDNase_nPk-processed", "rdmHistone_bPk-processed", "rdmHistone_gPk-imputed", "rdmHistone_gPk-processed", "rdmHistone_nPk-imputed", "rdmHistone_nPk-processed")

# Read in data in a loop
for (x in postfixes) {
  mtx <- gs_load_data(x)
  mtx <- mtx[, colnames(mtx) %in% term.mapping$Name]  # Subset by Names that are mapped
  mtx.cor <- Hmisc::rcorr(mtx, type=cortype)[[1]]
  # mtx.cor <- mine(mtx, n.cores=4)$MIC
  mtx.cor.long <- reshape_wide_data(mtx.cor)
  mtx.cor.epi.joined <- cbind(mtx.cor.epi.joined, mtx.cor.long[match(rownames(mtx.cor.epi.joined), rownames(mtx.cor.long)), ])
  colnames(mtx.cor.epi.joined)[length(colnames(mtx.cor.epi.joined))] <- paste("gs", x, sep=".")
}
```

```{r eval=FALSE}
# Visualize distributions
#mtx.cor.epi.joined <- normalizeQuantiles()
ggplot(reshape2::melt(mtx.cor.epi.joined), aes(x=Var2, y=value, fill=Var2)) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(-1, 1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))
```

Summarize all, and most correlated with episimilarity metrics. Visualize it as a heatmap

```{r similarityHeatmap_gwasCatalog, fig.height=9, fig.width=13, dev="png"}
#mtx.cor.epi.joined <- sweep(mtx.cor.epi.joined, 1, apply(mtx.cor.epi.joined, 1, median,na.rm=T))
mtx.cor.epi.joined <- scale(mtx.cor.epi.joined)

if (cortype != "MIC") {
  mtx.cor <- Hmisc::rcorr(as.matrix(mtx.cor.epi.joined), type=cortype)
  ## Clustering
  hclust.method <- "ward.D2"
  hclustergram <- hclust(as.dist(1-mtx.cor[[1]]), method=hclust.method) # Cluster it

  # Save the results
  write.xlsx2(formatC(mtx.cor[[1]], format="f", digits=2), "results/comparison_analysis.xlsx", sheetName=paste("corr", cortype, mat.type, sep="_"), append=TRUE)
  write.xlsx2(formatC(mtx.cor[[3]], format="e", digits=2), "results/comparison_analysis.xlsx", sheetName=paste("pval", cortype, mat.type, sep="_"), append=TRUE)
  
  # To have p-values overlayed on the cells: cellnote=formatC(mtx.summary[[1]], format="f", digits=2)
  # To have corr coeffs overlayed on the cells: cellnote=formatC(mtx.summary[[3]], format="e", digits=2)
  # Visualization only
  par(oma=c(7.3,0,0,5), mar=c(6, 4.1, 5.1, 5), cex.main=0.5) # Adjust margins
  gplots::heatmap.2(mtx.cor[[1]], Rowv=as.dendrogram(hclustergram), Colv=as.dendrogram(hclustergram), symbreaks = TRUE, trace="none", density.info="none", col=col3, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=1, cexCol=1, notecol="black", notecex=0.9, cellnote=formatC(mtx.cor[[1]], format="f", digits=2))
  
  # PDF only, overlay corr coeffs
  pdf(paste("results/comparison_analysis_corr_", cortype, "_", mat.type, ".pdf", sep=""), width=7.5, height=6)
  par(oma=c(6,0,0,6), mar=c(11, 4.1, 5.1, 6), cex.main=1) # Adjust margins
  gplots::heatmap.2(mtx.cor[[1]], Rowv=as.dendrogram(hclustergram), Colv=as.dendrogram(hclustergram), trace="none", density.info="none", col=col3, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=0.5, cexCol=0.5, notecol="black", notecex=0.25, cellnote=formatC(mtx.cor[[1]], format="f", digits=2))
  dev.off()
  # # PDF only, overlay pval coeffs
  # pdf(paste("results/comparison_analysis_pval_", cortype, ".pdf", sep=""), width=7.5, height=6)
  # par(oma=c(3.3,0,0,2.8), mar=c(11, 4.1, 5.1, 5), cex.main=0.5) # Adjust margins
  # heatmap.2(mtx.summary[[1]], trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=0.5, cexCol=0.6, notecol="black", notecex=0.25, cellnote=formatC(mtx.summary[[3]], format="e", digits=2))
  # dev.off()
} else {
  library(minerva) # Playing with MIC correlation coefficient
  dim(mtx.cor.epi.joined[ complete.cases(mtx.cor.epi.joined), ])
mtx.summary <- mine(mtx.cor.epi.joined[ complete.cases(mtx.cor.epi.joined), ], n.cores=4)$MIC
  # PDF only, overlay corr coeffs
  pdf(paste("results/comparison_analysis_corr_", cortype, ".pdf", sep=""), width=7.5, height=6)
  par(oma=c(3.3,0,0,2.8), mar=c(11, 4.1, 5.1, 5), cex.main=0.45) # Adjust margins
  gplots::heatmap.2(mtx.summary, trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=0.4, cexCol=0.4, notecol="black", notecex=0.4, cellnote=formatC(mtx.summary, format="f", digits=2), breaks=seq(-1, 1, 1, 20))
  dev.off()
}
```

Another view of the correlogram

```{r fig.height=8}
library(corrplot)
corrplot.mixed(mtx.cor[[1]], lower="color", upper="circle", order="original", col = col3(20), tl.pos = "lt", tl.cex=0.5, tl.col="black", cl.cex=0.5)
par(col="black")
pdf("results/comparison_analysis_corrplot.pdf", height = 20, width = 18)
corrplot(mtx.cor[[1]], method = "circle", type="upper", col=col3(20), tl.col = "black", tl.cex=1.4, cl.pos = NULL, diag=F)
dev.off()
```

Combining p-values

```{r eval=FALSE}
# https://stats.stackexchange.com/questions/59360/fishers-method-for-combing-p-values-what-about-the-lower-tail
pchisq( -2*sum(log(p-values)), df, lower.tail=FALSE)
df = 2*length(p-values)
```


Comparison with random chance
====

```{r similarityHeatmap_Random, fig.height=9, fig.width=13, dev="png", eval=FALSE}
# Random correlation matrix
set.seed(1)
mtx.summary <- rcorr(as.matrix(mtx.rand(as.matrix(mtx.cor.epi.joined), randomize="mix")) , type=cortype)
# Save the results
unlink("results/comparison_analysis.xlsx")
write.xlsx2(formatC(mtx.summary[[1]], format="f", digits=2), "results/comparison_analysis.xlsx", sheetName="corr_random", append=TRUE)
write.xlsx2(formatC(mtx.summary[[3]], format="e", digits=2), "results/comparison_analysis.xlsx", sheetName="pval_random", append=TRUE)

# To have p-values overlayed on the cells: cellnote=formatC(mtx.summary[[1]], format="f", digits=2)
# To have corr coeffs overlayed on the cells: cellnote=formatC(mtx.summary[[3]], format="e", digits=2)
pdf("results/comparison_analysis_corr_random.pdf", width=7.5, height=6)
par(oma=c(6,0,0,6), mar=c(11, 4.1, 5.1, 6), cex.main=1) # Adjust margins
heatmap.2(mtx.summary[[1]], trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=0.5, cexCol=0.6, notecol="black", notecex=0.3, cellnote=formatC(mtx.summary[[1]], format="f", digits=2))
dev.off()
# pdf("results/comparison_analysis_pval_random.pdf", width=7.5, height=6)
# par(oma=c(3.3,0,0,2.8), mar=c(11, 4.1, 5.1, 5), cex.main=0.5) # Adjust margins
# heatmap.2(mtx.summary[[1]], trace="none", density.info="none", col=color, distfun=function(x){dist(x, method=dist.method)}, hclustfun=function(x){hclust(x, method=hclust.method)}, cexRow=0.5, cexCol=0.6, notecol="black", notecex=0.25, cellnote=formatC(mtx.summary[[3]], format="e", digits=2))
# dev.off()
```

