---
title: Analysis of schizophrenia and bipolar disorder differentially methylated regions,
  compared with MWAS study
author: "Mikhail Dozmorov"
output: html_document
---

```{r setup, echo=FALSE, include=FALSE, cache=FALSE}
# Set up the environment
library(knitr) 
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', dpi=300, out.width=700, echo=F, autodep=T, width=200, comment=NA, warning=F, results='hide')
options(replace.assign=TRUE, width=500)
tidy.opts=list(width.cutoff=200)
suppressMessages(library(pander))
panderOptions('table.split.table', Inf)
set.seed(1)
```

```{r loadLibraries, warning=FALSE, message=FALSE}
source("../../R.GenomeRunner/utils1.R")
```

[Jerez, A., Rubin, L., Li, X., & Xu, C. (n.d.). Genome-wide methylome analyses reveal novel epigenetic regulation patterns in schizophrenia and bipolar disorder.](http://www.hindawi.com/journals/bmri/aa/201587/)

Supplementary tables 1 and 3 have genomic coordinates of  hyper/hypomethylated DMRs for schizophrenia and bipolar disorder, respectively. Note, coordinates are in hg18 genome assembly, and were lifted over to hg19 assembly. 

Number of DMRs:

- 128 bp_hyper.bed
- 198 bp_hypo.bed
- 351 sz_hyper.bed
- 602 sz_hypo.bed

Also included are [all MWAS results](http://www.people.vcu.edu/~ejvandenoord/Supplemental%20table%20-%20All%20MWAS%20results%20with%20q-value%20better%20than%200%2001.csv) with q-value better than 0.01 in “Methylome-Wide Association Study of Schizophrenia: Identifying Blood Biomarker Signatures of Environmental Insults. JAMA Psychiatry “ by Aberg KA, McClay JL, Nerella S, Clark S, Kumar G, Chen W, Khachane AN, Xie LY, Hudson A, Gao G, Harada A, Hultman CM, Sullivan PF, Magnusson PKE, van den Oord EJCG (2014), JAMA Psychiatry 2014; ePub Jan 8.

- 139 MWAS.bed

These MWAS do not overlap with the aforementioned DMRs.

**Hypothesis:** The disease-associated sets from the two studies will show similar regulatory enrichments.

These sets were tested for enrichment in 3 categories of regulatory datasets:

* **Histone** - 722 cell type-specific histone modification marks (41 marks, 69 cell lines). Information about individual marks can be looked up on [Histome: the Histone Infobase](http://www.actrec.gov.in/histome/searchgoog.php). For more detailed overview, check an excellent paper [Barski, A., Cuddapah, S., Cui, K., Roh, T.-Y., Schones, D. E., Wang, Z., … Zhao, K. (2007). High-resolution profiling of histone methylations in the human genome. Cell, 129(4), 823–37. doi:10.1016/j.cell.2007.05.009](http://www.sciencedirect.com/science/article/pii/S0092867407006009)

* **TFBSs** - 1952 cell type-specific transcription factor binding sites (277 transcription factors, 91 cell lines). Information about individual transcription factors can be looked up on [Factorbook](http://factorbook.org/mediawiki/index.php/Welcome_to_factorbook).

* **Roadmap Epigenomics** - histone modifications, brain specific subset

Information about cell lines used in the ENCODE project can be found at [the ENCODE cell types](http://genome.ucsc.edu/ENCODE/cellTypes.html) portal.

Information about Roadmap Epigenomics abbreviations can be found in the [jul2013.roadmapData.qc](https://docs.google.com/spreadsheet/ccc?key=0Am6FxqAtrFDwdHU1UC13ZUxKYy1XVEJPUzV6MEtQOXc&usp=sharing) document.

Instead of all available cell lines, subsets of tissue-specific cell lines can be used. 

* Brain cell lines, normal karyotype: c("Nha", "Hah", "Bcbrainh11058n")

The ENCODE datasets are cell-type incomplete, that is, for one cell type the data about the distribution of all histone marks are available, but the other cell type may have only a few histone mark datasets. So, the matrixes of *cells x marks* will have many missing values. They have to be filtered to remove rows/columns containing too much missing values, otherwise, clustering and visualization algorithms break. The heatmaps of such filtered matrixes are shown.

The [Roadmap Epigenomics](https://sites.google.com/site/anshulkundaje/projects/epigenomeroadmap) datasets are also incomplete, i.e., brain has only 7 histone modification marks profiled (h3k9me3, h3k4me3, h3k4me1, h3k36me3, h3k27me3, h3k9ac, h3k27ac).

P-values are corrected for multiple testing using FDR. For each analysis, only the significant enrichment results (p.adj < 0.1) are shown. Negative p-value indicates depletion.

Summary
===

BP and SCZ show enrichment in similar set of activating histone marks, some, like H2AZ, are bivalent, i.e., activating and repressing. These enrichments are observed in Normal human astrocytes.

MWAS was underpowered, but still able to show enrichment in activating H3K4me2 and H3K9ac marks.

H3K9ac (prefrontal cortex) and another activating mark H3K4me3 (germinal matrix) are enriched in BP. They are also enciched in SCZ across whole brain structures.

MWAS shows H3K4me1 (activating) in inferior temporal lobe, prefrontal cortex, hippocampus, and substantia nigra. Another activating H3K9ac (substantia nigra). Depletion in repressive marks H3K27me3 and H3K9me3.

```{r}
cells<-c("Nha", "Hah", "Bcbrainh11058n")
```

Results
===

Bipolar disorder
---

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/Histone/matrix.txt", colnum=c(7), factor="Histone", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=2, toPlot="bar")
```

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/TFBSs/matrix.txt", colnum=c(1), factor="Tfbs", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="bar")
```

Schizophrenia
---

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/Histone/matrix.txt", colnum=c(4), factor="Histone", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=2, toPlot="bar")
```

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/TFBSs/matrix.txt", colnum=c(5), factor="Tfbs", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=0, toPlot="bar")
```

MWAS
---

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/Histone/matrix.txt", colnum=c(3), factor="Histone", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=2, toPlot="bar")
```

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/TFBSs/matrix.txt", colnum=c(4), factor="Tfbs", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="bar")
```

Roadmap Epigenomics data
===

Bipolar disorder
---
Overview of Histone/Tissue enrichments

```{r fig.height=5 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(1), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=0, toPlot="heat")
```

Most significant enrichments/depletions

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(1), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=1, toPlot="bar")
```

Schizophrenia
---
Overview of Histone/Tissue enrichments

```{r fig.height=5 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(8), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=0, toPlot="heat")
```

Most significant enrichments/depletions

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(8), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=1, toPlot="bar")
```

MWAS
---

Overview of Histone/Tissue enrichments

```{r fig.height=5 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(4), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=0, toPlot="heat")
```

Most significant enrichments/depletions

```{r fig.height=3 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(4), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=1, toPlot="bar")
```



Differences between BP, SZ, and MWAS
===

Histone
---
```{r fig.height=3 , eval=T}
showHeatmap("data.gr/Histone/matrix.txt", colnum=c(7,4,3), factor="Histone", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=2, toPlot="bar")
```

TFBS
---
```{r fig.height=3 , eval=T}
showHeatmap("data.gr/TFBSs/matrix.txt", colnum=c(1,5,4), factor="Tfbs", cell=cells, isLog10=TRUE, adjust="fdr", pval=0.1, numtofilt=1, toPlot="bar")
```

Roadmap Epigenomics
---
```{r fig.height=3 , eval=T}
showHeatmap("data.gr/all_vs-reHistoneBrain/matrix.txt", colnum=c(1,8,4), factor="none", cell="none", isLog10=FALSE, adjust="none", pval=0.1, numtofilt=1, toPlot="bar")
```

